### 2018-05-06 paper/supplement.tex
Add plot to supplementary materials showing that posterior doesn't care if prior scale is doubled on hyperlocations

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf,float, bm, subcaption}
\graphicspath{C:/Users/Colin/Documents/GitHub/BB_data_analysis/paper/fig/}
\usepackage{enumerate}
\usepackage{natbib}
\bibliographystyle{asa}
\usepackage{url} % not crucial - just used below for the URL

% no longer require relative path to figs
\graphicspath{{fig/}}

\usepackage[x11names]{xcolor}
\usepackage{framed} % provides framed env., should be removed in final draft
\definecolor{shadecolor}{rgb}{1,.5,.5}

% math macros

\newcommand{\ind}{\stackrel{ind.}{\sim}}
\newcommand{\op}{\operatorname}
\DeclareMathOperator*{\argmin}{arg\,min}
\newcommand{\myequation}{\begin{equation}}
\newcommand{\myendequation}{\end{equation}}
\let\[\myequation
       \let\]\myendequation

\pdfminorversion=4
% NOTE: To produce blinded version, replace "0" with "1" below.
\newcommand{\blind}{0}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\begin{document}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

\begin{center}
{\large\bf SUPPLEMENTARY MATERIAL}
\end{center}

\begin{description}

\item Stan code for Model 4, estimated using the Backblaze data set (also available as a supplementary .csv file).

\end{description}

\section{Stan code for Model 4}
\label{sec:stan-code}
{\scriptsize
\begin{verbatim}
functions {
real sev_logpdf(real y, real mu, real sigma){
real z;
z = (y - mu) / sigma;
return -log(sigma) + z - exp(z);
}

real sev_logccdf(real y, real mu, real sigma){
return -exp((y - mu) / sigma);
}

real sev_logcdf(real y, real mu, real sigma){
return log1m_exp(-exp((y - mu) / sigma));
}
}

data {
int M;
int N_obs;
int N_cens;
real endtime_obs[N_obs];
real endtime_cens[N_cens];
real starttime_obs[N_obs];
real starttime_cens[N_cens];
int<lower=1, upper=M> dm_obs[N_obs];
int<lower=1, upper=M> dm_cens[N_cens];
vector<lower=0, upper=1>[2] p; # quantiles to model
}
transformed data{
vector[2] z_corr;
for(i in 1:2)
z_corr[i] = log(-1.0 * log1m(p[i])); # used to get location(mu) from quantile(t_p)
}
parameters{
real log_tp1;
real<lower=0> sigma1;
real eta_tp2;
real eta_s2;
real eta_pi;
real<lower=0> tau_tp2;
real<lower=0> tau_s2;
real<lower=0> tau_pi;
vector[M] log_tp2_raw;
real<lower=0, upper=1> sigma2[M];
vector[M] logit_pi_raw;
}

transformed parameters{
real mu1;
vector[M] mu2;
vector[M] log_pi;
mu1 = log_tp1 - sigma1 * z_corr[1];
for(m in 1:M){
mu2[m] = (eta_tp2 + tau_tp2 * log_tp2_raw[m]) - (sigma2[m] * z_corr[2]);
}
for(m in 1:M)
log_pi[m] = log_inv_logit(eta_pi + tau_pi * logit_pi_raw[m]);
}

model{
real tmp[2];
int m;
real logpi;
real mu_1;
real mu_2;
real sig_1;
real sig_2;

for(i in 1:N_obs){
m = dm_obs[i];
logpi = log_pi[m];
mu_2 = mu2[m];
sig_2 = sigma2[m];
// numerator:   log( p * f1 * (1 - F2) + f2 * (1 - p * F1) )
//            = log( exp(log(p) + log(f1) + log(1 - F2)) + exp(log(f2) + log(1 - exp(log(p) + log(F1)))) )
tmp[1] = log_sum_exp(logpi + sev_logpdf(endtime_obs[i], mu1, sigma1) +
sev_logccdf(endtime_obs[i], mu_2, sig_2),
sev_logpdf( endtime_obs[i], mu_2, sig_2) +
log1m_exp(logpi + sev_logcdf(endtime_obs[i], mu1, sigma1)
)
);
// denominator:  log((1 - p * F1) * (1 - F2))
//            =  log(1 - p * F1) + log(1 - F2)
tmp[2] = log1m_exp(logpi + sev_logcdf(starttime_obs[i], mu1, sigma1)) +
sev_logccdf(starttime_obs[i], mu_2, sig_2);

target += tmp[1] - tmp[2];
}

for(i in 1:N_cens){
m = dm_cens[i];
logpi = log_pi[m];
mu_2 = mu2[m];
sig_2 = sigma2[m];

// numerator:   log((1 - p * F1) * (1 - F2))
//            =  log(1 - p * F1) + log(1 - F2)
tmp[1] = log1m_exp(logpi + sev_logcdf(endtime_cens[i], mu1, sigma1)) +
sev_logccdf(endtime_cens[i], mu_2, sig_2);
// denominator:  log((1 - p * F1) * (1 - F2))
//            =  log(1 - p * F1) + log(1 - F2)
tmp[2] = log1m_exp(logpi + sev_logcdf(starttime_cens[i], mu1, sigma1)) +
sev_logccdf(starttime_cens[i], mu_2, sig_2);

target += tmp[1] - tmp[2];
}

log_tp1        ~ normal(7, 2);
log_tp2_raw    ~ normal(0, 1);
sigma1         ~ lognormal(0, 1);
sigma2         ~ lognormal(eta_s2, tau_s2);
logit_pi_raw   ~ normal(0, 1);
eta_tp2        ~ normal(9, 2);
eta_s2         ~ normal(0, 2);
eta_pi         ~ normal(-3, 1);
tau_tp2        ~ cauchy(0,1);
tau_s2         ~ cauchy(0,1);
tau_pi         ~ cauchy(0,1);
}

\end{verbatim}
}

\pagebreak
\section{Posterior predictive graphical checks}
Full set of Figures 5 and 6 for all 44 drive-models.  Numbering in the top left indicates drive-model.  Left: Adjusted K-M estimates of the proportion failing.  The original data (bold line) and 19 ``replicated" data sets from the posterior predictive distribution (dashed lines) are shown.  Right: Posterior median of proportion failing as a function of time for Models 1-4.  Plotted on Weibull probability scales.  The solid step function is the adjusted K-M estimate.  
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-v2-1.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-v2-2}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-v2-3.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-v2-4.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-v2-5.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-v2-6.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-v2-7.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-v2-8.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-v2-9.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-v2-10.pdf}
\end{figure}
\begin{figure}[H]
\includegraphics[width=\textwidth]{ppcheck-v2-11.pdf}
\end{figure}
\clearpage

\section{Density Plots for Prior Distributions}
To better visualize the prior distributions, below are plots of
the prior densities for Models 1 - 4.
\begin{figure}[H]
\center
\includegraphics[width=\textwidth]{priorsmod1.pdf}
\caption{Prior Densities for Model 1.  For $t_{p1 = 0.5}$, $t_{p2 = 0.2}$, $x$-axis is in thousands of hours.}
\end{figure}

\begin{figure}[H]
\center
\includegraphics[width=\textwidth]{priormod234.pdf}
\caption{Prior Densities for Models 2 - 4 (top left to bottom right). For $t_{p1 = 0.5}$, $t_{p2 = 0.2g}$, $x$-axis is in thousands of hours.}
\end{figure}

\section{Sensitivity check}
To check whether the posterior distribution of the location hyperparameters, $\eta_\pi$, $\eta_{t_{p_2}}$ and $\eta_{\sigma_2}$ in Model 4 was sensitive to our choice of prior distributions for these parameters, we refit Model 4 doubling the scale of the prior distributions, i.e.
$$\eta_\pi \sim \text{N}(-3,2^2),\; \eta_{t_{p_2}} \sim \text{N}(9,4^2),\; \eta_{\sigma_2} \sim \text{N}(0, 4^2).$$
Figure \ref{double_var} shows that the posterior was nearly identical with the wider prior distributions. This suggests that the posterior is overwhelmingly driven by the information from the data, rather than from the prior (for either choice of prior.)

\begin{figure}[H]
\center
\includegraphics[width=\textwidth]{double_var.pdf}
\caption{Comparison of prior and posterior for location hyperparameters, $\eta_\pi,\; \eta_{t_{p_2}}$ and $\eta_{\sigma_2}$ under the original priors from Section 6.1 and under a modified prior doubling the scale for these parameters.}
\label{double_var}
\end{figure}

\end{document}